---
title: "Arbeidskrav 3"
author: "Pia, Annlaug, Mari"
format: html
editor: visual
editor_options: 
  chunk_output_type: inline
---

```{r}
#| include: false

library(tidyverse)

set.seed(1)
population <- rnorm(1000000, mean = 1.5, sd = 3)


samp1 <- data.frame(y = sample(population, 8, replace = FALSE))

samp2 <- data.frame(y = sample(population, 40, replace = FALSE))


m1 <- lm(y ~ 1, data = samp1)
m2 <- lm(y ~ 1, data = samp2)

summary(m1)
summary(m2)
```

1.  Explain the estimate, SE, *t*-value, and *p*-value from the regression models that we created previously (`m1` and `m2`).

    Estimate m1 = 1,84, m2 = 1, 5642\
    SE m1= 1,251, m2 = 0,4774\
    t-value m1 = 1,47, m2 = 3,276\
    p-value m1 = 0,185, m2 = 0,00221\
    \
    Estimatet er gjennomsnittet i gruppene. Gjennomsnittet i gruppen med 8 samples er 1,84 og i gruppen med 40 samples er det 1,5642. \
    Standard error er en statistikk for å estimere variasjonen i en hypotetisk distribusjon av gjennomsnittet. Siden m1 har større SE har den større feilmargin dersom man skal gjøre testen på ny. T-verdi viser forskjellen mellom gjennomsnittene og om forskjellen mellom gruppene er tilfeldig. m2 har størst t-verdi som betyr at denne modellen skiller seg mer fra null-hypotesen enn m1. \
    p-verdien forteller sannynligheten for at null-hypotesen er sann. m2 hadde lavest p-verdi som er langt under grensen på 0,05, og kunne derfor være signifikant. m1 hadde høy p-verdi og ville blitt ikke-signifikant.

2.  Discuss what contributes to the different results in the two studies (`m1` and `m2`).

    Modellene har forskjellig gruppestørrelse. Større samples gir lavere standard error og lavere standardavvik.

3.  Why do we use the shaded area in the lower and upper tail of the *t*-distribution (See Figure \@ref(fig:t-dist-fig)).

    Vi ser på det skraverte onmrådet for å se om sannsynligheten for at gjennomsnittet i en sample avviker signifikant fra et kjent gjennomsnitt for populasjonen.

```{r}
#| include: false
# Create data frames to store the model estimates
results_8 <- data.frame(estimate = rep(NA, 1000), 
                      se = rep(NA, 1000), 
                      pval = rep(NA, 1000), 
                      n = 8)  

results_40 <- data.frame(estimate = rep(NA, 1000), 
                      se = rep(NA, 1000), 
                      pval = rep(NA, 1000), 
                      n = 40)

# A for loop used to sample 1000 studies, each iteration (i) will draw a new sample
# from the population. 

for(i in 1:1000) {
  
  # Draw a sample 
  samp1 <- data.frame(y = sample(population, 8, replace = FALSE))
  samp2 <- data.frame(y = sample(population, 40, replace = FALSE))

  # Model the data
  m1 <- lm(y ~ 1, data = samp1)
  m2 <- lm(y ~ 1, data = samp2)
  
  # Extract values from the models
  results_8[i, 1] <- coef(summary(m1))[1, 1]
  results_8[i, 2] <- coef(summary(m1))[1, 2]
  results_8[i, 3] <- coef(summary(m1))[1, 4]

  results_40[i, 1] <- coef(summary(m2))[1, 1]
  results_40[i, 2] <- coef(summary(m2))[1, 2]
  results_40[i, 3] <- coef(summary(m2))[1, 4]
  
  
}


# Save the results in a combined data frame

results <- bind_rows(results_8, results_40)

# Example code for copy and paste

# A two facets histogram can be created with ggplot2
results %>%
  ggplot(aes(pval)) + 
  geom_histogram() +
  facet_wrap(~ n)


# Count the proportion of tests below a certain p-value for each 
results %>%
  filter(pval < 0.05) %>%
  group_by(n) %>%
  summarise(sig_results = n()/1000)

# Using the pwr package
library(pwr)

pwr.t.test(n = 40, sig.level = 0.05, d = 1.5/3, type = "one.sample")
```

4.  Calculate the standard deviation of the `estimate` variable, and the average of the `se` variable for each of the study sample sizes (8 and 40). Explain why these numbers are very similar. How can you define the Standard Error (SE) in light of these calculations?

5.  Create a histogram (see example code below) of the *p*-values from each study sample-size. How do you interpret these histograms, what do they tell you about the effect of sample size on statistical power?

6.  Calculate the number of studies from each sample size that declare a statistical significant effect (specify a threshold for α, your significance level).

7.  Using the `pwr` package, calculate the power of a one-sample t-test, with a effect size of `1.5/3`, your specified significance level and sample sizes 8 and 40. Explain the results in the light of your simulations.
